---
title: "Location and Road Features"
author: "William Feng 13786355"
date: "14/09/2020"
output: html_document
---

# Getting the data directly from the GitHub Repo 
```{r, message=FALSE, warning=FALSE}
setwd("~/motorcycle_crashes")

library(readr)
locations_station <- read_csv("weather/locations_station.csv")
station_list <- read_csv("weather/station_list.csv")
monthly_avg <- read_csv("weather/monthly_avg.csv")
```

```{r, message=FALSE}
library(dplyr)
station_list <- filter(station_list, 
                       station_list$STA=='QLD') 

# Removing useless variables like ID columns.
station_list <- station_list[, -c(2:5, 8)]
monthly_avg <- monthly_avg[, -c(3)]
locations_station <- locations_station[, -c(13,14,18:21,23:25,27,28,47,49:53)]

# Merging weather station list with average monthly rainfall data. First, make the primary keys. 

monthly_avg <- monthly_avg %>% 
  rename('Site' = 'bom_id') 
station_rain <- full_join(station_list, 
                          monthly_avg, 
                          by = "Site")
station_rain <- na.omit(station_rain)

# Joining the data set by year, month and site name.
main <- full_join(locations_station, 
                  station_rain, 
                  by = c("Crash_Year" = "Year", 
                         "Crash_Month" = "Month_name",
                         "site_id_1" = "Site"))

main <- na.omit(main) 

# Filter data to only fatal accidents
fataldata <- filter(main, main$Crash_Severity=="Fatal")
```

# Exploratory Data Analysis

```{r, Proportion of crash severity}
severity <- as.data.frame(table(main$Crash_Severity))

ggplot(severity, aes(x = reorder(Var1, -Freq),
                 y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = c("#FFDDDD","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF"),
           colour="black") + 
  labs(x = "Crash Severity", 
       y = "Number of Accidents", 
       title = "Severity of Motorcycle Accidents", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/22757),1),"%"), 
                y = Freq), 
            vjust=-0.5, 
            size=3.5, 
            position = position_dodge(0.9)) + 
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000, 10000, 12000, 14000), limits = c(0,14000)) + 
  theme_bw()
```

# Main Crash Hour
```{r Crash hour time }
hour <- as.data.frame(table(main$Crash_Hour))

library(ggplot2)
ggplot(hour, aes(x = Var1, y = Freq, color=Var1)) + 
  geom_bar(stat = "identity", fill=c("#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#FFDDDD","#CCEEFF","#CCEEFF","#FFDDDD","#CCEEFF","#CCEEFF",	"#FFDDDD","#FFDDDD","#FFDDDD","#FFDDDD","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF",	"#CCEEFF", "#CCEEFF"), 
           colour="black") +
  labs(x = "Time of Day (hrs)", 
       y = "Number of Accidents", 
       title = "Motorcycle Accidents over 24 Hours", 
       subtitle = "(2001-2018). Red highlighted columns show increase in crashes", 
       caption = "Times are shown in 24-hour format") +
  geom_text(aes(label = Freq, 
                y = Freq), 
            vjust=-0.2, 
            size=3.5, 
            position = position_dodge(0.9),
            colour = "black") +
  theme_bw()
```
# Fatality only crash hour
```{r}
fatalhour <- as.data.frame(table(fataldata$Crash_Hour))

library(ggplot2)
ggplot(fatalhour, aes(x = Var1, y = Freq, color=Var1)) + 
  geom_bar(stat = "identity", fill=c("#FFDDDD","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF","#FFDDDD","#CCEEFF","#CCEEFF",	"#FFDDDD","#FFDDDD","#FFDDDD","#FFDDDD","#CCEEFF","#CCEEFF","#CCEEFF","#CCEEFF",	"#FFDDDD", "#CCEEFF"), 
           colour="black") +
  labs(x = "Time of Day (hrs)", 
       y = "Number of Fatal Accidents", 
       title = "Fatal Accidents over 24 Hours", 
       subtitle = "(2001-2018)", 
       caption = "Times are shown in 24-hour format") +
  geom_text(aes(label = Freq, 
                y = Freq), 
            vjust=-0.2, 
            size=3.5, 
            position = position_dodge(0.9),
            colour = "black") +
  theme_bw()
```
# main dataset by speed zone
```{r}
fatalspeed <- as.data.frame(table(fataldata$Crash_Speed_Limit))

library(ggplot2)
ggplot(fatalspeed, aes(x = Var1, y = Freq, color=Var1)) + 
 geom_bar(stat = "identity", 
           colour="black") +
  labs(x = "Time of Day (hrs)", 
       y = "Number of Fatal Accidents", 
       title = "Fatal Accidents over 24 Hours", 
       subtitle = "(2001-2018)", 
       caption = "Times are shown in 24-hour format") +
  geom_text(aes(label = Freq, 
                y = Freq), 
            vjust=-0.2, 
            size=3.5, 
            position = position_dodge(0.9),
            colour = "black") +
  theme_bw()
```

# Extracting road types (main)
```{r}
library(stringr)
substring_last <- str_extract_all(main$Crash_Street,"\\w+$") 
v <- as.character(substring_last)
sort(table(v)*100/22757, decreasing = TRUE)
```

Most common crash sites 
1. Road - 11288 = 49.60%
2. St - 5343 = 23.47$
3. Hwy - 2422 = 10.64%
4. Dr - 1183 = 5.19%
5. Ave - 623 = 2.73%
6. Tce - 208 = 0.91%
7. Pde - 178 = 0.78%
8. Ramp - 148 = 0.65%
9. Ct - 131 = 0.57%
10. Blvd - 127 = 0.55%

```{r Crash locations}
type <- as.data.frame(table(v))

type2 <- type[with(type,order(-Freq)),]
type2 <- type2[1:5,]

type2$v <- c("Road", "Street", "Highway", "Drive", "Avenue")

library(ggplot2)
ggplot(type2, aes(x = reorder(v, -Freq), 
                  y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF", 
           colour="black") +
  labs(x = "Road Type", 
       y = "Number of Accidents", 
       title = "Top 5 Most Common Motorcycle Crash Locations", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/22757),1),"%"), 
                y = Freq), 
            vjust=-0.5, 
            size=3.5, 
            position = position_dodge(0.9)) +
  scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000, 10000, 12000), limits = c(0,12000)) +
  theme_bw()
```
# Fatal only data

```{r}
library(stringr)
substring_last <- str_extract_all(fataldata$Crash_Street,"\\w+$") 
v2 <- as.character(substring_last)
sort(table(v2)*100/761, decreasing = TRUE)
```

Most common crash sites 
1. Road 59%
2. Hwy 16%
3. St 11%
4. Dr 4.6%
5. Ave 1.5%

```{r Crash locations}
type2 <- as.data.frame(table(v2))

type3 <- type[with(type,order(-Freq)),]
type3 <- type3[1:5,]

type3$v2 <- c("Road", "Highway", "Street", "Drive", "Avenue")

library(ggplot2)
ggplot(type3, aes(x = reorder(v2, -Freq), 
                  y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF", 
           colour="black") +
  labs(x = "Road Type", 
       y = "Number of Fatalities", 
       title = "Top 5 Most Common Motorcycle Crash Locations", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/761),1),"%"), 
                y = Freq), 
            vjust=-0.5, 
            size=3.5, 
            position = position_dodge(0.9)) +
  scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500), limits = c(0,500)) +
  theme_bw()
```

```{r Road feature type}
features <- as.data.frame(table(main$Crash_Roadway_Feature))

features$Var1 <- c("Bikeway","Bridge", "Forestry Road", "Five way", "Cross Intersection", "Interchange", "Multiple Road", "Roundabout", "T-Junction","Y-Junction", "Median Opening","Merge Lane", "No Feature", "Other", "Railway")

features2 <- features[with(features,order(-Freq)),]
features2 <- features2[1:5,]

ggplot(features2, 
       aes(x = reorder(Var1, -Freq), 
           y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF",
           colour="black") +
  labs(x = "Road Feature", 
       y = "Number of Accidents", 
       title = "Top 5 Most Common Motorcycle Crash Road Features", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/22757),1),"%"), 
                y = Freq), 
            vjust=-0.5, 
            size=3.5, 
            position = position_dodge(1)) +
    scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000,10000, 12000), limits = c(0,12500)) +
  theme_bw()
```

```{r}
features <- as.data.frame(table(main$Crash_Roadway_Feature))

features$Var1 <- c("Bikeway","Bridge", "Forestry Road", "Five way", "Cross Intersection", "Interchange", "Multiple Road", "Roundabout", "T-Junction","Y-Junction", "Median Opening","Merge Lane", "No Feature", "Other", "Railway")

features2 <- features[with(features,order(-Freq)),]
features2 <- features2[1:5,]

ggplot(features2, 
       aes(x = reorder(Var1, -Freq), 
           y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF",
           colour="black") +
  labs(x = "Road Feature", 
       y = "Number of Accidents", 
       title = "Top 5 Most Common Motorcycle Crash Road Features", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/22757),1),"%"), 
                y = Freq), 
            vjust=-0.5, 
            size=3.5, 
            position = position_dodge(1)) +
    scale_y_continuous(breaks = c(0, 2000, 4000, 6000, 8000,10000, 12000), limits = c(0,12500)) +
  theme_bw()
```
# fatality only featureless roads
```{r}
library(ggplot2)
features2 <- as.data.frame(table(fataldata$Crash_Roadway_Feature))

features2$Var1 <- c("Bikeway","Bridge", "Forestry Road", "Cross Intersection", "Interchange", "Roundabout", "T-Junction", "Median Opening","Merge Lane", "No Feature", "Railway")

features3 <- features2[with(features2,order(-Freq)),]
features3 <- features3[1:5,]

ggplot(features3, 
       aes(x = reorder(Var1, -Freq), 
           y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF",
           colour="black") +
  labs(x = "Road Feature", 
       y = "Number of Fatal Accidents", 
       title = "Fatal accidents by road feature", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/761),1),"%"), 
                y = Freq), 
            vjust=-0.5, 
            size=3.5, 
            position = position_dodge(1)) +
    scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600), limits = c(0,600)) +
  theme_bw()
```


# Crash nature 
```{r, include = FALSE}
nature <- as.data.frame(table(main$Crash_Nature))

library(ggplot2)
ggplot(nature, aes(x = reorder(Var1, Freq), 
                 y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF", 
           colour="black") +
  labs(x = "Nature of Crash", 
       y = "Number of Accidents", 
       title = "Crash Nature", 
       subtitle = "Based on data from 2001-2018") +
  geom_text(aes(label = Freq, 
                y = Freq), 
            vjust=-0.2, 
            size=3.5, 
            position = position_dodge(0.9)) +
  theme_bw() +
  coord_flip()
```


```{r Crash group description, include=FALSE, eval=FALSE}
description <- as.data.frame(table(main$Crash_DCA_Group_Description))

description2 <- description[with(description,order(-Freq)),]
description2 <- description2[1:5,]

library(ggplot2)
ggplot(description2, aes(x = reorder(Var1, Freq), 
                 y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF", 
           colour="black") +
  labs(x = "Reason of Crash", 
       y = "Number of Accidents", 
       title = "Top 5 Causes of Crashes", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/22757),1),"%"), 
                y = Freq), 
            hjust=-0.1, 
            size=3.5, 
            position = position_dodge(1)) +
  scale_y_continuous(limits = c(0,3500)) + 
  theme_bw() +
  coord_flip()
```

# Leaflet map - heat map 

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(leaflet)
library(leaflet.extras)
library(RColorBrewer)

motorcycle <- awesomeIcons(
  icon = 'motorcycle',
  iconColor = 'orange',
  markerColor = "black",
  library = 'fa'
)

m <- leaflet(main) %>%
  addProviderTiles(providers$Stamen.Toner, group = "Black and White") %>%
  
  addTiles(options = providerTileOptions(noWrap = TRUE), group="Colour") %>%
  
  addMiniMap(tiles = providers$Stamen.Toner) %>%
  
  addHeatmap(
    lng = ~Crash_Longitude_GDA94,
    lat = ~Crash_Latitude_GDA94,
    radius = 15,
    blur = 25,
    gradient = "Oranges",
    cellSize = 25,
    intensity = 7,
    group = "Fatalities2"
) %>%

 addLayersControl(
   overlayGroups = c("Accidents"),
   baseGroups = c("Black and White", "Colour"),
   options = layersControlOptions(collapsed = FALSE)
 )
```

```{r}
m
```


# Featureless Roads Investigation 
```{r}
inv <- filter(main, main$Crash_Roadway_Feature == "No Roadway Feature")

inv2 <- as.data.frame(table(inv$Crash_DCA_Group_Description))

featurelessreason <- inv2[with(inv2,order(-Freq)),]
featurelessreason <- featurelessreason[1:10,]

ggplot(featurelessreason, aes(x = reorder(Var1, -Freq), 
                 y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF", 
           colour="black") +
  labs(x = "Reason of Crash", 
       y = "Number of Accidents", 
       title = "Top 10 Causes of Crashes on Featureless Roads", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/22757),1),"%"),  
                y = Freq), 
            vjust=-0.5,
            size=3.5, 
            position = position_dodge(0.5)) +
   scale_y_continuous(limits = c(0,2000)) +
  theme_bw() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

# fatal data only
```{r}
inv2 <- filter(fataldata, fataldata$Crash_Roadway_Feature == "No Roadway Feature")

inv3 <- as.data.frame(table(inv2$Crash_DCA_Group_Description))

featurelessreason2 <- inv3[with(inv3,order(-Freq)),]
featurelessreason2 <- featurelessreason2[1:11,]

library(ggplot2)
ggplot(featurelessreason2, aes(x = reorder(Var1, -Freq), 
                 y = Freq)) + 
  geom_bar(stat = "identity", 
           fill = "#CCEEFF", 
           colour="black") +
  labs(x = "Reason of Crash", 
       y = "Number of Accidents", 
       title = "Top 10 Causes of Crashes on Featureless Roads", 
       subtitle = "(2001-2018)") +
  geom_text(aes(label = paste(round((Freq*100/761),1),"%"),  
                y = Freq), 
            vjust=-0.5,
            #hjust=-0.1,
            size=3.5, 
            position = position_dodge(0.5)) +
   scale_y_continuous(limits = c(0,200)) +
  theme_bw() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

