---
title: "Refined R code"
author: "William Feng"
date: "04/09/2020"
output: html_document
---

# NOTES TO READ B4 STARTING
# CURRENTLY THERE'S AN ERROR there is a problem when we joined the rainfall data to the main set

-----------------------------------------------

# Getting data directly from the GitHub
```{r}
library(readr)
locations_station <- read_csv("weather/locations_station.csv")
station_list <- read_csv("weather/station_list.csv")
monthly_avg <- read_csv("weather/monthly_avg.csv")
```

Current state of data 
GPS coords - found in station_list
Rainfall amount - monthly_avg
Distance calculations & fatality data - locations_station

# Prepping data for the merging
1) Station list - extract site ID and GPS coords for QLD only
```{r, message=FALSE}
library(dplyr)
station_list <- filter(station_list, station_list$STA=='QLD') 
station_list <- station_list[, -c(2:5, 8)]
```

2) Monthly Avg
```{r}
monthly_avg <- monthly_avg[, -c(3)]
```

3) Location_station - already cleaned on GitHub (Showing only QLD and motorcycles > 0) - see #-ed part for raw dataset
```{r}
locations_station <- locations_station[, -c(13,14,18:21,23:25,27,28,47,49:53)]
```


# Primary Key dataset merging 

station_list PK is Site
monthly_avg PK is bom_id

```{r}
# merging station_list with monthly avg first
colnames(monthly_avg)
monthly_avg <- monthly_avg %>% rename('Site' = 'bom_id') 
```
Interim/proxy table
```{r}
station_rain <- full_join(station_list, monthly_avg, by = "Site")
station_rain <- na.omit(station_rain)
```

Merge proxy table with main locations_station dataset
```{r}
main <- full_join(locations_station, station_rain, by = c("Crash_Year" = "Year", "Crash_Month" = "Month_name", "site_id_1" = "Site"))
# joins the data set by year, month and site name. 
```

```{r}
main <- na.omit(main) # filters all NAs produced by the join 
```

```{r save dataset to rds file}
# save the data frame as an Rds object, reload with: df <- read_rds("main.Rds")
write_rds(main, "main.Rds")
```
# Exploratory 
```{r}
str(main)
```
# data spans 2001 - 2018, defines scope of the dataset 
```{r}
summary(main$Crash_Year) 
```
# find number of crashes per year and plot on simple graph 
```{r, message=FALSE}
library(sqldf)
year_count = sqldf("SELECT DISTINCT COUNT(Crash_Year) FROM main GROUP BY Crash_Year")
```

```{r}
table(main$Crash_Year)
```

# Plot crash year on x axis vs count on y axis 
```{r}
Year = c(2001:2018)
Count = unlist(year_count, use.names=FALSE)
plot(Year, Count)
Crash = cbind(Year, Count)
Crash = as.data.frame(Crash)
```
Prettify the above graph using ggplot2
```{r, message=FALSE}
library(ggplot2)
ggplot(Crash, aes(x=Year, y=Count)) + 
  geom_point() + 
  geom_smooth() + 
  scale_x_continuous(breaks=c(2001:2018), labels=c(2001:2018)) +
  labs(x = "Year", y= "Count", title = "Number of Motorcycle Crashes by Year")
```

# ERROR there is a problem when we joined the rainfall data to the main set

# Feasibility check - Does the average|total rainfall of the month increase fatalities? 
```{r}
#summary(main$rainfall) 
```

```{r}
table(main$Crash_Month, main$Crash_Year)
```


```{r}
#onth_count <- sqldf("SELECT DISTINCT COUNT(Crash_Month), SUM(rainfall), Crash_Year, Crash_Month FROM main WHERE Crash_Year = 2002 GROUP BY Crash_Month")
#ggplot(main, aes())
```


# First attempt at Leaflet location plotting 
## with crash coords layer + weather station coords layer + (future) chloropeth average rainfall 
```{r}
library(leaflet)
library(leaflet.extras)
library(leaflet.providers)
leaflet(main) %>% 
  addProviderTiles(providers$Stamen.Toner, group = "Black and white") %>% 
  
  addTiles(options = providerTileOptions(noWrap = TRUE), group="Colour") %>% 
  
  addMarkers(
    lng = ~Crash_Longitude_GDA94, 
    lat = ~Crash_Latitude_GDA94,
    clusterOptions = markerClusterOptions(),
    label = ~Crash_Severity,
    group = "Crash"
  ) %>%
  addMarkers(
    lng = ~Lon, 
    lat = ~Lat,
    clusterOptions = markerClusterOptions(),
    label = ~site_name_1,
    group = "Weather Station"
  ) %>%
  addLayersControl(
    baseGroups = c("Black and white","Colour"),
    overlayGroups = c("Crash", "Weather Station"),
    options = layersControlOptions(collapsed = FALSE)
  )

```


