---
title: "Workbook to investigate Question 1"
author: "Paul Gittings"
date: "09/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages and libraries}

# function to load library, will first attempt to load package if not present
setupPackage <- function( packageName, loud=FALSE ) {
  if (!require(packageName, character.only=TRUE)) {
    install.packages(packageName, verbose = loud, dependencies = TRUE)
  } 
  library(packageName, character.only=TRUE, verbose = loud)
}

setupPackage("tidyverse")
setupPackage("fastDummies")
setupPackage("PerformanceAnalytics")
setupPackage("corrr")
setupPackage("magrittr")
#setupPackage("polr")  #  ordinal logistic regression - not available for R 4.0.2 :-(, polr also in MASS package
setupPackage("ordinal")  # ordinal logistic regression

```

# Introduciton

The question to be investigated is: Does a relationship exist between the nature of the crash and the crash severity in motorcycle accidents?

Load the main crash data set and assign it to crashDF. EDA should reference crashDF

```{r load dataset}
  crashDF <- read_rds("../datasets/main.Rds")
```

# EDA

```{r get column names}
col_names <- names(crashDF)
print(col_names)
```
```{r get attribute values}
unique(crashDF$Crash_Severity)
#TODO:: turn Crash_Severity into an ordinal - property damage only, Minor Injury, Medical treatment, Hospitalisation, Fatal

unique(crashDF$Crash_Nature)

#NOTE: Crash_Sevrity ~ Crash_Nature is predicting an ordinal factor based on a nominal one
```
```{r historgram crash severity by carsh nature}

ggplot(crashDF, aes(x=Crash_Severity)) +
  geom_bar(stat="count") +
  facet_wrap(~Crash_Nature) +
  theme(axis.text.x = element_text(angle = 90))

```

```{r total accidents by crash nature}

ggplot(crashDF) +
    geom_bar(aes(x = forcats::fct_infreq(Crash_Nature), fill = Crash_Severity)) +
    theme(axis.text.x = element_text(angle = 90)) +
    xlab("Crash Nature")
```
Looking at the above plot, would it make sense to roll oup some of the Crash_Nature levels, say from "Hit Parked Vehicle" and all those to it's left?

```{r total accidents by crash severity}

ggplot(crashDF) +
    geom_bar(aes(x = forcats::fct_infreq(Crash_Severity), fill = Crash_Nature)) +
    theme(axis.text.x = element_text(angle = 90)) +
    xlab("Crash Severity")
```

# Engineer Data

Steps taken:

* Created Crash_Severity_Fac  which is an ordinal factor version of Crash_Severity
* Created Crash_Severity_Num which is a numeric version of Crash_Severity_Fac
* created dummy attributes for Crash_Nature
* all columnm names renamed to replace spaces with "_".
* created non_fatal_accident column, equals 1 for a non-fatal accident, 0 for a fatal accident
* created Crash_Nature_Num as a numeric version of Crash_Nature



NOTE: from this point on analysis should use the q1CrashDF  dataframe

```{r engineer variables}
# will check to see if we previously saved away the engineered data set and load it 
# rather than re-engineering the main crash data again.
# NOTE: this means that if additional engineering code is added the saved copy of q1CrashDF.Rds
# should be deleted OR the forceDfRebuild variable should be set to TRUE.
forceDfRebuild <- TRUE
q1DfFileName <-"q1CrashDF.Rds"

if( (!forceDfRebuild) & file.exists(q1DfFileName)) {
  q1CrashDF <- read_rds(q1DfFileName)
} else {
  
  
  # creat extra dummy columns for the Crash_Nature - should we drop the first one? 
  #     remove_first_dummy = TRUE
  q1CrashDF <- fastDummies::dummy_cols(crashDF, select_columns= "Crash_Nature")
  names(q1CrashDF)<-str_replace_all(names(q1CrashDF), c(" " = "_" , "," = "", "-"="_" ))
  names(q1CrashDF)<-str_replace_all(names(q1CrashDF), c("___" = "_"  ))
  
  # replace spaces with underscore in Crash_Severity
  q1CrashDF$Crash_Severity <- str_replace_all(q1CrashDF$Crash_Severity, " ", "_")
  
  # may not need this as we have Count_Fatal, Count_Hospitalisation etc 
  q1CrashDF$Crash_Severity_Fac <- factor(q1CrashDF$Crash_Severity, c("Property_damage_only",
        "Minor_injury",  "Medical_treatment", "Hospitalisation", "Fatal"), ordered=TRUE)
  
  # numeric version of Crash_Severity_Fac
  q1CrashDF$Crash_Severity_Num <- as.numeric(q1CrashDF$Crash_Severity_Fac)
  
  q1CrashDF %<>% mutate(non_fatal_accident = Count_Casualty_Fatality==0)
  
  q1CrashDF$Crash_Nature_Num <- as.numeric(factor(q1CrashDF$Crash_Nature))
  print(names(q1CrashDF))
   
  write_rds(q1CrashDF, q1DfFileName)
  
}
  

```

# Additional EDA

```{r correlation matrix 1 - all}
my_data <- select(q1CrashDF,  Count_Casualty_Fatality, Count_Casualty_Hospitalised,
                  Count_Casualty_MedicallyTreated, Count_Casualty_MinorInjury,
                  Crash_Nature_Angle, Crash_Nature_Collision_miscellaneous,
                  Crash_Nature_Fall_from_vehicle, Crash_Nature_Head_on,
                  Crash_Nature_Hit_animal, Crash_Nature_Hit_object,
                  Crash_Nature_Hit_parked_vehicle, Crash_Nature_Hit_pedestrian,
                  Crash_Nature_Non_collision_miscellaneous, Crash_Nature_Overturned,
                  Crash_Nature_Rear_end, Crash_Nature_Sideswipe,
                  Crash_Nature_Struck_by_external_load
                  )

#chart.Correlation(my_data, histogram=TRUE, pch=19)
```


```{r correlattion matrix 2: fatalities}
my_data <- select(q1CrashDF,  Count_Casualty_Fatality, 
                  Crash_Nature_Angle, Crash_Nature_Collision_miscellaneous,
                  Crash_Nature_Fall_from_vehicle, Crash_Nature_Head_on,
                  Crash_Nature_Hit_animal, Crash_Nature_Hit_object,
                  Crash_Nature_Hit_parked_vehicle, Crash_Nature_Hit_pedestrian,
                  Crash_Nature_Non_collision_miscellaneous, Crash_Nature_Overturned,
                  Crash_Nature_Rear_end, Crash_Nature_Sideswipe,
                  Crash_Nature_Struck_by_external_load
                  )
#chart.Correlation(my_data, histogram=TRUE, pch=19)
```
Above correlation matrixes are too small, but the results don't look that good. Let's try a text based approach to get the correltions.

```{r alterative to correlation graphs}
corrData <- select(q1CrashDF,  Count_Casualty_Fatality, Count_Casualty_Hospitalised,
                  Count_Casualty_MedicallyTreated, Count_Casualty_MinorInjury,
                  Crash_Nature_Angle, Crash_Nature_Collision_miscellaneous,
                  Crash_Nature_Fall_from_vehicle, Crash_Nature_Head_on,
                  Crash_Nature_Hit_animal, Crash_Nature_Hit_object,
                  Crash_Nature_Hit_parked_vehicle, Crash_Nature_Hit_pedestrian,
                  Crash_Nature_Non_collision_miscellaneous, Crash_Nature_Overturned,
                  Crash_Nature_Rear_end, Crash_Nature_Sideswipe,
                  Crash_Nature_Struck_by_external_load
                  )
 result <- corrData %>% correlate()
 
  
 result %>% focus(names(result)[2])  # Can't seem to be able to use Count_Casulaty_Fatality directly?
 
result %>% focus(names(result)[3])
 
result %>% focus(names(result)[4])

result %>% focus(names(result)[5])
  

```

Correlation analysis doesn't reveal any correlation of any strength between severity of accident and the nature of the accident. The Highest positive correlation is 0.12 between Head On accidents and Fatalities, however this i not a significant correlation. All remaining correlations are above -0.01 and below 0.01  which are no correlation at all.


Go back to the engineering level and introduce a new variable non_fatal_accidents and see if that correlation's exist for it. 

```{r correlation for non-fatal accidents}
corrData <- select(q1CrashDF,  non_fatal_accident,
                  Crash_Nature_Angle, Crash_Nature_Collision_miscellaneous,
                  Crash_Nature_Fall_from_vehicle, Crash_Nature_Head_on,
                  Crash_Nature_Hit_animal, Crash_Nature_Hit_object,
                  Crash_Nature_Hit_parked_vehicle, Crash_Nature_Hit_pedestrian,
                  Crash_Nature_Non_collision_miscellaneous, Crash_Nature_Overturned,
                  Crash_Nature_Rear_end, Crash_Nature_Sideswipe,
                  Crash_Nature_Struck_by_external_load
                  )
 result <- corrData %>% correlate()
 
 result %>% focus(names(result)[2])
```

Correlation still low. It would seem that the nature of an accident by itself has no correlation to the crash severity. However as can seen by the histograms at the start of the document a large number of accidents are the result of specific types of accidents, namely a collision at an angle and riders falling off. It is not clear if riders fall off as a result of some other accident type.


Just in case I'm doing things wrong with the above correlation let's look at a chi-squared test

```{r chi-squared test}

chisq.test(q1CrashDF$Crash_Severity_Fac, q1CrashDF$Crash_Nature, correct=FALSE)

```

Given the low p value the chi-squared test does imply that the Nature and Severity are dependent.

```{r quick glm model}


#  split data into 70% train 30% test.
trainSize <- floor(0.7*nrow(q1CrashDF))
set.seed(42)

trainIndexes <- sample(seq_len(nrow(q1CrashDF)), size = trainSize)
trainDF  <- q1CrashDF[trainIndexes, ]
testDF   <- q1CrashDF[-trainIndexes, ]

# model 1
# glm - gaussian
# dependent varialbe - numeric version of Crash Severity
# independent variables - all dummy variables of Crash_Nature
model1 <- glm(data=trainDF, Crash_Severity_Num ~ Crash_Nature_Angle +  Crash_Nature_Collision_miscellaneous +
             Crash_Nature_Fall_from_vehicle + Crash_Nature_Head_on +
                  Crash_Nature_Hit_animal + Crash_Nature_Hit_object +
                  Crash_Nature_Hit_parked_vehicle+ Crash_Nature_Hit_pedestrian +
                  Crash_Nature_Non_collision_miscellaneous +Crash_Nature_Overturned +
                  Crash_Nature_Rear_end + Crash_Nature_Sideswipe +
                  Crash_Nature_Struck_by_external_load, family="gaussian")

summary(model1)
```

Used a multiple liear regression model, looking at the summary it doesn't look that promising because all p-values for independent are over .05.

```{r get predictions}

# prediciton throws up a "rank-deficient fit may be misleading" warning
# this is a catch all type of warning, that could inlcude rank (in this case the data has the same rank),
# sparse data or other concerns predict has about the data,  it should be thought of as 
# "predicitons are brittle
#  https://stackoverflow.com/a/49862494
predictions <- predict(model1, newdata=testDF, type="response")

 print(paste("Model 1 accuracy:", sum(round(predictions)==testDF$Crash_Severity_Num)/nrow(testDF)))
```

Model is not very poor on accuracy.

Try agin but remove crash nature with low occurrence

```{r model 2}
# model 2
# glm - gaussian
# dependent varialbe - numeric version of Crash Severity
# independent variables - selected (highest incidence) dummy variables of Crash_Nature
model2<- glm(data=trainDF, Crash_Severity_Num ~ Crash_Nature_Angle +  
             Crash_Nature_Fall_from_vehicle + Crash_Nature_Head_on +
                  Crash_Nature_Hit_animal + Crash_Nature_Hit_object +
                  Crash_Nature_Hit_parked_vehicle+ Crash_Nature_Hit_pedestrian +
                  Crash_Nature_Rear_end + Crash_Nature_Sideswipe, family="gaussian")

summary(model2)
predictions <- predict(model2, newdata=testDF, type="response")

 print(paste("Model 2 accuracy:", sum(round(predictions)==testDF$Crash_Severity_Num)/nrow(testDF)))
```

Very slight improvement in AIC so model is a slightly better fit but prediciton accuracy reamins unchanged.


Let's back up a bit. Rather than using dummy variables for the Crash_Nature, lets use the numeric re-encoded column in the next model

```{r model 3}
# model 3
# glm - gaussian
# dependent varialbe - numeric version of Crash Severity
# independent variables - numeric version  Crash_Nature - does this make sense as Crash_Nature is non-ordered
model3 <- glm(data=trainDF, Crash_Severity_Num ~ Crash_Nature_Num, family="gaussian")

summary(model3)
predictions <- predict(model3, newdata=testDF, type="response")

 print(paste("Model 3 accuracy:", sum(round(predictions)==testDF$Crash_Severity_Num)/nrow(testDF)))
 
```

Accuracy even worse.

# Ordinal Logistic Regression

Would have liked to have used the polr package but not available for this version of R, so will use the ordinal package and it's function for cumulative link models for ordinal regresion (clm).
Reference: https://marissabarlaz.github.io/portfolio/ols/

Not really expecting a different result to the attempted linear regression, just including for completeness.

Predictions are all hospitalisation!

# TODO play with this a bit more.

```{r Oridnal Logistic Regression model1}

model4 <- clm(data=trainDF, Crash_Severity_Fac ~ Crash_Nature_Angle +  Crash_Nature_Collision_miscellaneous +
             Crash_Nature_Fall_from_vehicle + Crash_Nature_Head_on +
                  Crash_Nature_Hit_animal + Crash_Nature_Hit_object +
                  Crash_Nature_Hit_parked_vehicle+ Crash_Nature_Hit_pedestrian +
                  Crash_Nature_Non_collision_miscellaneous +Crash_Nature_Overturned +
                  Crash_Nature_Rear_end + Crash_Nature_Sideswipe +
                  Crash_Nature_Struck_by_external_load, link = "logit" )
summary(model4)

predictions <- predict(model4, newdata=testDF, type="class")

print(paste("Model 3 accuracy:", sum(as.numeric(predictions$fit)==as.numeric(testDF$Crash_Severity_Fac))/nrow(testDF)))
```
The result looks better at .55 but not great. However, it seems all the predictions are "Hospitalisation" which suggests something is not working.
